---
title: "In-class Exercise 2"
author: "Ha Duc Tien"
date: "September 02, 2024"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  message: false
  freeze: true
---

# 1. Getting started

```{r}
pacman::p_load(sf, raster, spatstat, tmap, tidyverse, maptools)
```

# 2. Importing the spatial data

```{r}
childcare_sf <- st_read("data/child-care-services-geojson.geojson") %>%
  st_transform(crs = 3414)
```

```{r}
sg_sf <- st_read(dsn = "data", layer="CostalOutline")%>%
  st_transform(crs = 3414)
```

```{r}
mpsz_sf <- st_read(dsn = "data", 
                layer = "MP14_SUBZONE_WEB_PL")%>%
  st_transform(crs = 3414)
```
# 3. Some insights and guidance

Working with st_union()

The code chunk below, st_union() is used to derive the coastal outline of sf tibble data.frame

```{r}
sg_sf <- mpsz_sf %>%
  st_union()
```

sg_sf will look similar to the code below

```{r}
plot(sg_sf)
```

Since maptools expired, we could use the code below instead of `as.SpatialGridDataFrame.im()`

```{r, fig.width = 12, fig.height = 9, dpi = 100}
#| eval: false
gridded_kde_childcareSG_bw <- as(kde_childcareSG.bw, "SpatialGridDataFrame")
plot(gridded_kde_childcareSG_bw)
spplot(gridded_kde_childcareSG_bw)
```

For Monte Carlo simulation use

```{r}
set.seed(1234)
```

Use Edge correction method of spatstat to reduce biases of spatial graphs

# 4. Take-home Exercise 1

The study area is Bangkok Metropolitan Region

:::callout-tip

The WGS84/UTM zone 47N the EPSG code is 32647

:::

The data Thai traffic accident 2019 - 2022

Use the code below to import the data

```{r}
accident_data <- read_csv("data/thai_road_accident_2019_2022.csv")%>%
  filter(!is.na(longitude) & longitude != "",
         !is.na(latitude) & latitude != "") %>%
  st_as_sf(coords = c("longitude", "latitude"), 
           crs = 4326) %>%
  st_transform(crs = 32647)
```