---
title: "In-class Exercise 10"
author: "Ha Duc Tien"
date: "October 21, 2024"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  message: false
  freeze: true
---

# 1. Getting started

```{r}
pacman::p_load(tidyverse, sf, tmap, httr, performance)
```

# 2. Importing data

```{r}
#| eval: false
file_list <- list.files(path = "data/aspatial",
                        pattern = "^ResidentialTransaction.*\\.csv$",
                        full.names = TRUE)

realis_data <- file_list %>%
  map_dfr(read_csv)
```

# 3. Data wrangling

The code chunk below does the following

-   converting values in Sale Date field from character to numerical date format
-   extracting resale and condominium transaction records.

```{r}
#| eval: false
condo_resale <- realis_data %>%
  mutate(`Sale Date` = dmy(`Sale Date`)) %>%
  filter(`Type of Sale` == "Resale" &
           `Property Type` == "Condominium")
```

Preparing data

```{r}
#| eval: false

postcode <- unique(condo_resale$`Postal Code`)
```


Geocoding

```{r}
#| eval: false

url <- "https://onemap.gov.sg/api/common/elastic/search"
found <- data.frame()
not_found <- data.frame()

for (postcode in postcode){
  query <- list('searchVal'=postcode, 'returnGeom'='Y', 
                'getAddrDetails'='Y', 'pageNum'='1')
  res <- GET(url, query=query)
  if ((content(res)$found)!=0){
    found <- rbind(found, data.frame(content(res))[4:13])
  } else {not_found = data.frame(postcode)
  }
}
```

Tidying field names

```{r}
#| eval: false

found <- found %>%
  select(c(6:8)) %>%
  rename(POSTAL = `results.POSTAL`,
         XCOORD = `results.X`,
         YCOORD = `results.Y`)
```

Write a code chunk to join condo_resale and found. Name the output condo_resale_geocoded.

```{r}
#| eval: false

condo_resale_geocoded = left_join(
  condo_resale, found, 
  by = c('Postal Code' = 'POSTAL'))
```

Write a code chunk to convert condo_resale_geocoded from tibble data frame to sf point feature data frame.

```{r}
#| eval: false

condo_resale_sf <- st_as_sf(condo_resale_geocoded, 
                            coords = c("XCOORD",
                                       "YCOORD"),
                            crs=3414)
```

The code chunk below is used to check if there are overlapping point features.

```{r}
#| eval: false

overlapping_points <- condo_resale_sf %>%
  mutate(overlap = lengths(st_equals(., .)) > 1)
```

In the code code chunk below, st_jitter() of sf package is used to move the point features by 5m to avoid overlapping point features.

```{r}
#| eval: false

condo_resale_sf <- condo_resale_sf %>%
  st_jitter(amount = 2)
```

```{r}
#| eval: false

write_rds(condo_resale_sf, "data/rds/condo_resale_sf.rds")
```

```{r}
condo_resale_sf <- read_rds("data/rds/condo_resale_sf.rds")

condo_resale_sf
```